Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})

### 2014-05-17 manuscript.tex
edits to results, tables, figures cont.

The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}. The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations. If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Fig. \ref{Fig3Structure}).

@@

We performed a second grow-out  of teosinte to assess whether lighting conditions or sample size may have affected our ability to detect and effect of \emph{tb1}.  For the second grow-out we measured tillering index every five days through day 50 for 302 plants. We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele. We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p<0.0611), with no significant correlation on any day. Similarly, relationships between \emph{Hopscotch} genotype and tiller number, and \emph{Hopscotch} genotype and culm diameter are not significant.

@@

We find no evidence of recent introgression in our analyses. Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.

 % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
\begin{table}[htbp]
  \centering
  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
    \begin{tabular}{ccccc}
  %  \toprule
          & \multicolumn{2}{c}{\emph{tb1} region} & \multicolumn{2}{c}{Chr 1} \\
 %   \midrule
    Population & Maize assignment & Teosinte assignment & Maize assignment & Teosinte assignment \\
    Ejutla A & 0.02158681 & 0.9784132 & 0.2026814 & 0.7973186 \\
    Ejutla B & 0.01888194 & 0.9811181 & 0.1872131 & 0.8127869 \\
    La Mesa & 0.0118675 & 0.9881333 & 0.8068998 & 0.1931017 \\
    San Lorenzo & 0.01551389 & 0.9844861 & 0.2048252 & 0.7951748 \\
 %   \bottomrule
    \end{tabular}
  \label{Table4Q}
\end{table}

>>>>

The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}. The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations. If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig. \ref{Fig3Structure}).

 % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering
  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
    \begin{tabular}{ccccc} \
          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
    \end{tabular}
  \label{Table4Q}
\end{table}

@@

We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.  For the second grow-out we measured tillering index every five days through day 50 for 302 plants. We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele. We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p$<$0.0611), with no significant correlation on any day. Similarly, relationships between \emph{Hopscotch} genotype and tiller number and \emph{Hopscotch} genotype and culm diameter were not significant.

@@

We find no evidence of recent introgression in our analyses. Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, Table \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.
